PANCHANGE_data_preprocessing

### 2020-08-11 PANCHANGE_data_preprocessing2020-07-15.Rmd
if statement added for pcl prepan

```{r glad prepandemic pcl cleaning}
if(GLAD == TRUE) {
glad.pcl.raw.id <- glad.pcl.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    pcl.stressful_experience_repeated_images_prepan = pcl.stressful_experience_repeated_images,
    pcl.stressful_experience_upset_reminded_prepan = pcl.stressful_experience_upset_reminded,
    pcl.stressful_situation_avoiding_activities_prepan = pcl.stressful_situation_avoiding_activities,
    pcl.cut_people_feeling_distant_prepan = pcl.cut_people_feeling_distant,
    pcl.feeling_irritable_or_having_angry_outbursts_prepan = pcl.feeling_irritable_or_having_angry_outbursts,
    pcl.difficulty_concentrating_prepan = pcl.difficulty_concentrating
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

*Note that scores with missingness will be changed to NA later in the script
```{r pcl sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
pcl.items_key_prepan <- c(1,1,1,1,1,1)

@@

# Look at values
prepan.pcl.raw.id %>%
  descr(pcl.sum_score_prepan_raw,
        stats = "common"
        )
```

@@

prepandemic
```{r pcl prepan missingness}
glad.pcl.raw.id <- glad.pcl.raw.id %>%
  mutate(
    na_per_person_pcl_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[pcl.items_prepan])
      )
  )

glad.pcl.raw.id %>%
  freq(na_per_person_pcl_prepan,
       cumul = F)
```

```{r pcl prepan sum score if missigng items set to NA}
glad.pcl.raw.id <- glad.pcl.raw.id %>%
  mutate(
    pcl.sum_score_prepan =
      if_else(condition = na_per_person_pcl_prepan > 0,
              true = NA_real_,
              false = pcl.sum_score_prepan_raw,
              missing = NA_real_
      )
  )

glad.pcl.raw.id %>%
  descr(pcl.sum_score_prepan_raw, stats = "common")

glad.pcl.raw.id %>%
  descr(pcl.sum_score_prepan, stats = "common")
```

```{r glad pcl head}
head(glad.pcl.raw.id)
```

```{r glad pcl summary}
summarytools::dfSummary(
  glad.pcl.raw.id,
  graph.col = F,
  valid.col = F, labels.col = F)
```

@@

```{r glad phq cleaning}
glad.phq.raw.id <- glad.phq.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    phq.little_interest_or_pleasure_in_doing_things_prepan = phq.little_interest_or_pleasure_in_doing_things,
    phq.feeling_down_depressed_or_hopeless_prepan = phq.feeling_down_depressed_or_hopeless,
    phq.staying_asleep_sleeping_trouble_prepan = phq.staying_asleep_sleeping_trouble,
    phq.feeling_tired_or_having_little_energy_prepan = phq.feeling_tired_or_having_little_energy,
    phq.poor_appetite_or_overeating_prepan = phq.poor_appetite_or_overeating,
    phq.feeling_bad_failure_family_prepan = phq.feeling_bad_failure_family,
    phq.trouble_concentrating_reading_newspaper_prepan = phq.trouble_concentrating_reading_newspaper,
    phq.moving_fidgety_noticed_opposite_prepan = phq.moving_fidgety_noticed_opposite,
    phq.dead_hurting_thoughts_prepan = phq.dead_hurting_thoughts
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.phq.raw.id)
# Inspect colnames
colnames(glad.phq.raw.id)
#Differences
dim(glad.phq.raw)[1]-dim(glad.phq.raw.id)[1]
```

>>>>

```{r glad prepandemic pcl cleaning}
if(GLAD == TRUE) {
exclude_cols <- c("Sample",
                  "ID")
glad.pcl.raw.id <- glad.pcl.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    pcl.stressful_experience_repeated_images_prepan = pcl.stressful_experience_repeated_images,
    pcl.stressful_experience_upset_reminded_prepan = pcl.stressful_experience_upset_reminded,
    pcl.stressful_situation_avoiding_activities_prepan = pcl.stressful_situation_avoiding_activities,
    pcl.cut_people_feeling_distant_prepan = pcl.cut_people_feeling_distant,
    pcl.feeling_irritable_or_having_angry_outbursts_prepan = pcl.feeling_irritable_or_having_angry_outbursts,
    pcl.difficulty_concentrating_prepan = pcl.difficulty_concentrating
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

*Note that scores with missingness will be changed to NA later in the script
```{r pcl sum score prepan}
if(GLAD == TRUE | EDGI == TRUE){
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
pcl.items_key_prepan <- c(1,1,1,1,1,1)

@@

# Look at values
prepan.pcl.raw.id %>%
  descr(pcl.sum_score_prepan_raw,
        stats = "common"
        )
}
```

@@

prepandemic
```{r pcl prepan missingness}
if(GLAD == TRUE | EDGI == TRUE){
prepan.pcl.raw.id <- prepan.pcl.raw.id %>%
  mutate(
    na_per_person_pcl_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[pcl.items_prepan])
      )
  )

prepan.pcl.raw.id %>%
  freq(na_per_person_pcl_prepan,
       cumul = F)
}
```

```{r pcl prepan sum score if missigng items set to NA}
if(GLAD == TRUE | EDGI == TRUE){
prepan.pcl.raw.id <- prepan.pcl.raw.id %>%
  mutate(
    pcl.sum_score_prepan =
      if_else(condition = na_per_person_pcl_prepan > 0,
              true = NA_real_,
              false = pcl.sum_score_prepan_raw,
              missing = NA_real_
      )
  )

prepan.pcl.raw.id %>%
  descr(pcl.sum_score_prepan_raw, stats = "common")

prepan.pcl.raw.id %>%
  descr(pcl.sum_score_prepan, stats = "common")
}
```

```{r glad pcl head}
if(GLAD == TRUE | EDGI == TRUE){
head(prepan.pcl.raw.id)
}
```

```{r glad pcl summary}
if(GLAD == TRUE | EDGI == TRUE){
summarytools::dfSummary(
  prepan.pcl.raw.id,
  graph.col = F,
  valid.col = F, labels.col = F)
}
```

@@

```{r glad phq cleaning}
if(GLAD == TRUE){
glad.phq.raw.id <- glad.phq.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    phq.little_interest_or_pleasure_in_doing_things_prepan = phq.little_interest_or_pleasure_in_doing_things,
    phq.feeling_down_depressed_or_hopeless_prepan = phq.feeling_down_depressed_or_hopeless,
    phq.staying_asleep_sleeping_trouble_prepan = phq.staying_asleep_sleeping_trouble,
    phq.feeling_tired_or_having_little_energy_prepan = phq.feeling_tired_or_having_little_energy,
    phq.poor_appetite_or_overeating_prepan = phq.poor_appetite_or_overeating,
    phq.feeling_bad_failure_family_prepan = phq.feeling_bad_failure_family,
    phq.trouble_concentrating_reading_newspaper_prepan = phq.trouble_concentrating_reading_newspaper,
    phq.moving_fidgety_noticed_opposite_prepan = phq.moving_fidgety_noticed_opposite,
    phq.dead_hurting_thoughts_prepan = phq.dead_hurting_thoughts
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.phq.raw.id)
# Inspect colnames
colnames(glad.phq.raw.id)
#Differences
dim(glad.phq.raw)[1]-dim(glad.phq.raw.id)[1]
}
```
